---
title: "1 - microcolony production figures and analysis"
author: "Michelle Z Hotchkiss"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

#Packages

```{r}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(MetBrewer)
library(lme4) 
library(lmerTest)
library(performance) 
library(emmeans)
library(glmmTMB)
```

#Load and format data

```{r}
mcdata <- read_csv("data/mc_production.csv")

mcdata$treatment <- case_match(mcdata$treatment, "Control" ~ "Control", "DMSO" ~ "DMSO", "1ppb" ~ "1 ng/mL", "10ppb" ~ "10 ng/mL", "100ppb" ~ "100 ng/mL", "800ppb" ~ "800 ng/mL")

mcdata$treatment <- factor(mcdata$treatment, levels = c("Control", "DMSO", "1 ng/mL", "10 ng/mL", "100 ng/mL", "800 ng/mL"))

mcdata$sc_id <- as.factor(mcdata$sc_id)
mcdata$age <- as.factor(mcdata$age)

mcdata$average_pupal_weight <- (mcdata$pupae_weight/mcdata$pupae_count)
```

#Figures

##1A: Number of adults alive

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1A <- 
  mcdata %>% 
  ggplot(aes(x=treatment, y=worker_count, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age), labeller=age_labeller)+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
     strip.text = element_text(colour = "black"),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Number of workers", limits = c(3.9,10.5))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```

##1B: Adult weight

```{r}
# New facet label names for age variable
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1B<-
  mcdata %>% 
  ggplot(aes(x=treatment, y=average_worker_weight, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age), labeller=age_labeller)+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
    strip.text = element_text(colour = "black"),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Average worker weight (mg)", limits = c(120,221))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```


##1C: Number of eggs

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1C<-
  mcdata %>% 
  ggplot(aes(x=treatment, y=egg_count, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
     strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Number of eggs", limits = c(0,59))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```

##1D: Egg weight

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1D<-
mcdata %>% 
  ggplot(aes(x=treatment, y=average_egg_weight, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
     strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Average egg weight (mg)", limits = c(1.1,2.6))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```

##1E: Number of larvae

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1E<-
mcdata %>% 
  ggplot(aes(x=treatment, y=larvae_count, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
    strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Number of larvae", limits = c(0, 41))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```


##1F: Larval weight

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1F<-
mcdata %>% 
  ggplot(aes(x=treatment, y=average_larvae_weight, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
    strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_blank(),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Average larva weight (mg)", limits = c(0, 285))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```


##1G: Number of pupae

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1G<-
mcdata %>% 
  ggplot(aes(x=treatment, y=pupae_count, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
    strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_text(size = 16, color = "black", angle = 45, hjust = 1, vjust=1),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Number of pupae", limits = c(0, 3.4))+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```


##1H: Pupal weight

```{r}
age_names <- list(
  '1'="Age: 1 Week",
  '2'="Age: 3 Weeks"
)

age_labeller <- function(variable,value){
  return(age_names[value])
}

mcprod_fig1H<-
mcdata %>% 
  ggplot(aes(x=treatment, y=average_pupal_weight, fill=treatment, color=treatment)) + 
  geom_boxplot(alpha=0.4, lwd=0.7)+
  geom_jitter(shape=16, position=position_jitter(width=0.15, height=0), size=3)+
  facet_grid(cols=vars(age))+
  theme_classic()+
  theme(
    panel.spacing = unit(.20, "lines"),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1), 
    strip.background = element_rect(
     color="white", linewidth=1.5, linetype="solid"),
    strip.text = element_blank(),
    axis.line=element_blank(),
    text = element_text(size = 16), axis.text.x = element_text(size = 16, color = "black", angle = 45, hjust = 1, vjust=1),
        axis.text.y = element_text(size = 16, color = "black"),
        legend.position = "none",
        axis.title.y = element_text(size = 16, margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 16, margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  scale_x_discrete(name="Treatment")+
  scale_y_continuous(name ="Average pupa \nweight (mg)")+
  scale_fill_met_d("Hiroshige", direction = -1)+
  scale_color_met_d("Hiroshige", direction = -1)
```

##Multi-panel Figure

```{r}
mcprod_final<-
  ggarrange(mcprod_fig1A + rremove("xlab"),
            NULL,
          mcprod_fig1B + rremove("xlab"),
          mcprod_fig1C+ rremove("xlab"),
          NULL,
          mcprod_fig1D + rremove("xlab"), 
          mcprod_fig1E + rremove("xlab"), 
          NULL,
          mcprod_fig1F + rremove("xlab"), 
          mcprod_fig1G + rremove("xlab"),
          NULL,
          mcprod_fig1H + rremove("xlab"), 
          labels = c("A", "", "B", "C", "", "D", "E", "", "F", "G", "","H"),
          label.x = 0.01,
          label.y = 0.98,
          font.label = list(size = 16),
          ncol = 3, nrow = 4,
          widths = c(1,0.05,1),
          heights = c(1,1,1,1),
          align = "v")

annotate_figure(mcprod_final,
                    bottom = text_grob("Treatment", size=16))
```


#Analysis

##1A: Number of adults


```{r}
adultcountA <- glmer(worker_count ~ treatment*age + (1|sc_id), data = mcdata, family = poisson)

adultcountB <- glmer(worker_count ~ treatment + age +(1|sc_id), data = mcdata, family = poisson)

anova(adultcountA, adultcountB)

adultcountC <- glmer(worker_count ~ treatment + (1|sc_id), data = mcdata, family = poisson)

anova(adultcountB,adultcountC)

adultcountD <- glmer(worker_count ~ age + (1|sc_id), data = mcdata, family = poisson)

anova(adultcountB,adultcountD)

check_model(adultcountB)
performance::check_overdispersion(adultcountB)
```


##1B: Adult weight

```{r}
adultsizeA <- lmer(average_worker_weight ~ treatment*age + (1|sc_id), data = mcdata)

anova(adultsizeA)

adultsizeB <- lmer(average_worker_weight ~ treatment + age + (1|sc_id), data = mcdata)

anova(adultsizeB)
ranova(adultsizeB)

check_model(adultsizeB)
```

##1C: Number of eggs

```{r}
eggcountA <- glmer(egg_count ~ treatment*age + (1|sc_id), data = mcdata, family = poisson)

eggcountB <- glmer(egg_count ~ treatment+age + (1|sc_id), data = mcdata, family = poisson)

anova(eggcountA, eggcountB)

check_model(eggcountA)
performance::check_overdispersion(eggcountA)
#overdispersion detected - try a negative binomial distribution

eggcountA_nb <- glmer.nb(egg_count ~ treatment*age + (1|sc_id), data = mcdata)
#model fails to converge, try using glmmTMB 

#according to https://stats.stackexchange.com/questions/419304/computing-different-types-of-negative-binomial-regression, glmer.nb uses similar negative binomial formula to nbinom1

eggcountA_TMB <- glmmTMB(egg_count ~ treatment*age + (1|sc_id), data = mcdata, family = nbinom1())

eggcountB_TMB <- glmmTMB(egg_count ~ treatment+age + (1|sc_id), data = mcdata, family = nbinom1())

anova(eggcountA_TMB,eggcountB_TMB)
#interaction not signficant, return to glmer.nb for consistency

eggcountB_nb <- glmer.nb(egg_count ~ treatment+age + (1|sc_id), data = mcdata)

eggcountC_nb <- glmer.nb(egg_count ~ age + (1|sc_id), data = mcdata)

eggcountD_nb <- glmer.nb(egg_count ~ treatment + (1|sc_id), data = mcdata)

anova(eggcountB_nb, eggcountC_nb)
anova(eggcountB_nb, eggcountD_nb)

#nothing significant

check_model(eggcountB_nb)
performance::check_overdispersion(eggcountB_nb)
```



##1D: Egg weight

```{r}
eggweightA <- lmer(average_egg_weight ~ treatment*age + (1|sc_id), data = mcdata)

eggweightB <- lmer(average_egg_weight ~ treatment+age + (1|sc_id), data = mcdata)

anova(eggweightA, eggweightB)

eggweightC <- lmer(average_egg_weight ~ treatment + (1|sc_id), data = mcdata)

eggweightD <- lmer(average_egg_weight ~ age + (1|sc_id), data = mcdata)

anova(eggweightB, eggweightC)
anova(eggweightB, eggweightD)

check_model(eggweightB)

```

##1E: Number of larvae

```{r}
larvaecountA <- glmer(live_larvae ~ treatment*age + (1|sc_id), data = mcdata, family = poisson)

larvaecountB <- glmer(live_larvae ~ treatment+age + (1|sc_id), data = mcdata, family = poisson)

anova(larvaecountA, larvaecountB)

larvaecountC <- glm(live_larvae ~ age + (1|sc_id), data = mcdata, family = poisson)

larvaecountD <- glmer(live_larvae ~ treatment + (1|sc_id), data = mcdata, family = poisson)

anova(larvaecountB, larvaecountC)
anova(larvaecountB, larvaecountD)

check_model(larvaecountB)
performance::check_overdispersion(larvaecountB)
#overdispersion detected, try negative binomial distribution


larvaecountA_nb <- glmer.nb(live_larvae ~ treatment*age + (1|sc_id), data = mcdata)

larvaecountB_nb <- glmer.nb(live_larvae ~ treatment+age + (1|sc_id), data = mcdata)

anova(larvaecountA_nb,larvaecountB_nb)

larvaecountC_nb <- glmer.nb(live_larvae ~ treatment + (1|sc_id), data = mcdata)

larvaecountD_nb <- glmer.nb(live_larvae ~ age + (1|sc_id), data = mcdata)

anova(larvaecountB_nb,larvaecountC_nb)
anova(larvaecountB_nb,larvaecountD_nb)

check_model(larvaecountD_nb)
performance::check_overdispersion(larvaecountD_nb)

larvaecountD_nb_means <- emmeans(larvaecountD_nb, specs = pairwise ~ age)
larvaecountD_nb_means
```

##1F: Larval Weight

```{r}
larvalweightA <- lmer(average_larvae_weight ~ treatment*age + (1|sc_id), data = mcdata)

larvalweightB <- lmer(average_larvae_weight ~ treatment+age + (1|sc_id), data = mcdata)

anova(larvalweightA, larvalweightB)

check_model(larvalweightA)

tukey<-TukeyHSD(larvalweightA)

larvalweightA_means = emmeans(larvalweightA, specs = ~ treatment*age)
larvalweightA_means

#compare means with custom list of contrasts

Con1 <- c(1,0,0,0,0,0,0,0,0,0,0,0)
DMSO1 <- c(0,1,0,0,0,0,0,0,0,0,0,0)
CHT11 <- c(0,0,1,0,0,0,0,0,0,0,0,0)
CHT101 <- c(0,0,0,1,0,0,0,0,0,0,0,0)
CHT1001 <- c(0,0,0,0,1,0,0,0,0,0,0,0)
CHT8001 <- c(0,0,0,0,0,1,0,0,0,0,0,0)
Con3 <- c(0,0,0,0,0,0,1,0,0,0,0,0)
DMSO3 <- c(0,0,0,0,0,0,0,1,0,0,0,0)
CHT13 <- c(0,0,0,0,0,0,0,0,1,0,0,0)
CHT103 <- c(0,0,0,0,0,0,0,0,0,1,0,0)
CHT1003 <- c(0,0,0,0,0,0,0,0,0,0,1,0)
CHT8003 <- c(0,0,0,0,0,0,0,0,0,0,0,1)

contrast(larvalweightA_means, method = list("Con1 - DMSO1" = Con1 - DMSO1,
                                            "Con1 - CHT11" = Con1 - CHT11,
                                            "Con1 - CHT101" = Con1 - CHT101,
                                            "Con1 - CHT1001" = Con1 - CHT1001,
                                            "Con1 - CHT8001" = Con1 - CHT8001,
                                            "DMSO1 - CHT11" = DMSO1 - CHT11,
                                            "DMSO1 - CHT101" = DMSO1 - CHT101,
                                            "DMSO1 - CHT1001" = DMSO1 - CHT1001,
                                            "DMSO1 - CHT8001" = DMSO1 - CHT8001,
                                            "CHT11 - CHT101" = CHT11 - CHT101,
                                            "CHT11 - CHT1001" = CHT11 - CHT1001,
                                            "CHT11 - CHT8001" = CHT11 - CHT8001,
                                            "CHT101 - CHT1001" = CHT101 - CHT1001,
                                            "CHT101 - CHT8001" = CHT101 - CHT8001,
                                            "CHT1001 - CHT8001" = CHT1001 - CHT8001,
                                            "Con3 - DMSO3" = Con3 - DMSO3,
                                            "Con3 - CHT13" = Con3 - CHT13,
                                            "Con3 - CHT103" = Con3 - CHT103,
                                            "Con3 - CHT1003" = Con3 - CHT1003,
                                            "Con3 - CHT8003" = Con3 - CHT8003,
                                            "DMSO3 - CHT13" = DMSO3 - CHT13,
                                            "DMSO3 - CHT103" = DMSO3 - CHT103,
                                            "DMSO3 - CHT1003" = DMSO3 - CHT1003,
                                            "DMSO3 - CHT8003" = DMSO3 - CHT8003,
                                            "CHT13 - CHT103" = CHT13 - CHT103,
                                            "CHT13 - CHT1003" = CHT13 - CHT1003,
                                            "CHT13 - CHT8003" = CHT13 - CHT8003,
                                            "CHT103 - CHT1003" = CHT103 - CHT1003,
                                            "CHT103 - CHT8003" = CHT103 - CHT8003,
                                            "CHT1003 - CHT8003" = CHT1003 - CHT8003,
                                            "Con1 - Con3" = Con1 - Con3,
                                            "DMSO1 - DMSO3" = DMSO1 - DMSO3,
                                            "CHT11 - CHT13" = CHT11 - CHT13,
                                            "CHT101 - CHT103" = CHT101 - CHT103,
                                            "CHT1001 - CHT1003" = CHT1001 - CHT1003,
                                            "CHT8001 - CHT8003" = CHT8001 - CHT8003
                                            ))

```

##1G: Number of pupae

```{r}
pupaecountA <- glmer(pupae_count ~ treatment*age + (1|sc_id), data = mcdata, family = poisson)

pupaecountB <- glmer(pupae_count ~ treatment+age + (1|sc_id), data = mcdata, family = poisson)

anova(pupaecountA, pupaecountB)

pupaecountC <- glmer(pupae_count ~ age + (1|sc_id), data = mcdata, family = poisson)

pupaecountD <- glmer(pupae_count ~ treatment + (1|sc_id), data = mcdata, family = poisson)

anova(pupaecountB, pupaecountC)
anova(pupaecountB, pupaecountD)
#age is significant

check_model(pupaecountC)
performance::check_overdispersion(pupaecountC)
#overdispersion, try nb


pupaecountA_nb <- glmer.nb(pupae_count ~ treatment + age + (1|sc_id), data = mcdata)

pupaecountB_nb <- glmer.nb(pupae_count ~ age + (1|sc_id), data = mcdata)

pupaecountC_nb <- glmer.nb(pupae_count ~ treatment + (1|sc_id), data = mcdata)

anova(pupaecountA_nb, pupaecountB_nb)
anova(pupaecountA_nb, pupaecountC_nb)

check_model(pupaecountB_nb)
performance::check_overdispersion(pupaecountB_nb)

pupaecountB_nb_means <- emmeans(pupaecountB_nb, specs = pairwise ~ age)
pupaecountB_nb_means
```

##1H: Pupal weight

Too little data for accurate analysis.



